DeepSphere: a graph-based spherical CNN

### 2019-09-19 main.tex
Updates from Overleaf

% The plan for this section
% 0.1 Explain the sampling problem
\paragraph{Sampling} 
As the graph convolution act on a sampled version of the signal $\b{f}$ and the rotation on its continuous counterpart $f$, we first need to define a sampling operator.
Given a sampling scheme $V=\{v_i\in\mathbb S^2, i=0, \dots, n\}$ of the sphere and a function $f : \mathbb S^2 \supset F_V \to \mathbb R$, the sampling operator $T_V: L^2(\mathbb S^2) \supset F\to \mathbb R^n,\  T_V(f) = \b{f}$ is defined as $\b{f}:\ f_i=f(v_i)$, where $F_V$ suitable subspace of $L^2(\mathbb S^2)$ such that $T_V$ is invertible, i.e., we can unambiguously reconstruct the function $f\in F_V$ from its sampled values $\b{f}$.
The existence of such subspace depends on the sampling scheme $V$, and its characterization is a common problem in signal processing \cite{Driscoll:1994:CFT:184069.184073}. 
For most samplings, it is not known if $F_V$ exist and hence if $T_V$ is invertible. A special case is the \textit{equiangular sampling} scheme where a sampling theorem holds, and thus we have a closed form of $T_V^{-1}$ is known. % \cite{Driscoll:1994:CFT:184069.184073}
For sampling schemes where no such sampling formula is available, we leverage the discrete SHT to reconstruct $f$ from $\bf{f}$, thus approximating $T_V^{-1}$. 
For all theoretical consideration, we assume that $F_V$ exists and $f \in F_V$.

% 0.2 Explain why the Laplacian is so important...
\paragraph{Problem formulation} 
By definition, our (spherical) graph convolution is rotation equivariant, if and only if, it commutes with the rotation operator defined as $R(g), g\in SO(3)$: $R(g) f(x) = f\left(g^{-1} x \right)$. 
In the context of this work, graph convolution is performed by recursive applications of the graph Laplacian, i.e: $h(\b{L}) \b{f} = \left(\sum_{i=0}^K \alpha_i \b{L}^i\right) \b{f}$\nati{Put this equation somewhere else?}.  Hence, if $\b{L}$ commutes with the rotation operator, then, by recursion, it will also commute with the convolution $h(\b{L})$.
As a result, to simplify our analysis, we concentrate on the graph Laplacian and $h(\b{L})$ is rotational invariant if and only if 
% 0.3 Formulate the problem
\begin{equation} \label{eq:equivariance}
	\b{R}_V(g) \b{L} \b{f} = \b{L} \b{R}_V(g) \b{f} \hspace{1cm} \forall f\in F_V \text{ and } \forall g\in SO(3), 
\end{equation}
where $\b{R}_V(g) = T_V R(g) T_V^{-1}$. For practical application, we define the \textit{normalized equivariance error} for a signal $\b{f}$, and a  rotation $g$ as:
\begin{equation} \label{eq:equivariance error}
	E_{\b{L}}(\b{f}, g) = \left(\frac{ \norm {\b{R}_V(g) \b{L} \b{f} - \b{L} \b{R}_V(g) \b{f}} }{\norm {\b{f}}}\right)^2,
\end{equation}
More generally for a class of signal $f \in C \subset F_V$, the \textit{mean equivariance error}
\begin{equation} \label{eq:mean equivariance error}
	\overline E_{\b{L}, C} = \mathbb E_{\b{f}\in C, g\in SO(3)} \ E_{\b{L}}(\b{f}, g).
\end{equation}
In practice, to anlyse the equivariance error frequency,  the set $C$ we used is composed of functions $f$ with made of spherical harmonic of a single degree $\ell$.
The expected value is obtained by averaging over a finite number of random functions and random rotations.

@@

% 2) explain our approach
We chose a different approach and take inspiration from \cite{belkin2005towards}, which proves that for a random \emph{uniform sampling} scheme the graph Laplacian converges toward the Laplace-Beltrami operator $\Delta$ as the number of sampling point goes to infinity.
This convergence result is important as a) $\Delta$ commutes with rotation and b) is diagonalized by the spherical harmonics. 
In this case the weighting scheme is full (every node is connected to every node) and is based the exponential kernel: 
\nati{@martino, please check this equation... Do you use 4t?}
\begin{equation} \label{eq:belkin_weights}
w_{ij} = e^{-\frac{1}{t} \norm{x_i-x_j}^2}
\end{equation}
In practice
The following theorem extend the work of Belkin and Nyiogi to fixed sampling \nati{(Including Healpix or not?)}. 
Given a sampling $V = \{x_0, \dots, x_{n-1}\}$ define $\sigma_i$ to be the patch of the surface of the sphere corresponding to $x_i$, define $A_i$ to be its corresponding area and $d_i$ to be the radius of the smallest ball in $\mathbb R^3$ containing $\sigma_i$. Define $d^{(n)} := \max_{i=0, \dots, n}d_i$ and $A^{(n)}=\max_{i=0, \dots, n}A_i$.

\begin{theorem}
	For a sampling $\mathcal P = \{x_i\in\mathbb S^2\}_{i=0}^{n-1}$ of the sphere that is equi area and such that $d^{(n)} \leq \frac{C}{\sqrt{n}}$, for all $f: \mathbb S^2 \rightarrow \mathbb R$ Lipschitz with respect to the euclidean distance in $\mathbb R^3$, for all $y\in\mathbb S^2$, there exists a sequence $t_n = n^\beta$ such that the rescaled Heat Kernel Graph Laplacian operator $\frac{|\mathbb S^2|}{4\pi t_n}L^t_n$ converges pointwise to the Laplace Beltrami operator on the sphere $\triangle_{\mathbb S^2}$  for $n\to\infty$:
	$$ \lim_{n\to\infty}\frac{|\mathbb S^2|}{4\pi t_n} L_n^{t_n}f(y) =  \triangle_{\mathbb S^2}f(y).$$
	\label{theo:pointwise convergence for a regular sampling}
\end{theorem}

This is \emph{not} a proof of equivariance of the graph convolution for a specific sampling. We simply observe that the graph Laplacian converges pointwisely toward an operator that commutes with rotation. To have the desired equality in eq..., we would need to prove a stronger result, for example uniform convergence (in the space of the operators).
\nati{Should we put this here?: }
This does not hold in general \cite{belkin2007convergence}. However, we believe that this is achievable for a bandlimited signals.
More importantly, the proof of theorem \ref{theo:pointwise convergence for a regular sampling} in Appendix ... inspires us in the construction of our graph Laplacian. a) use gaussian kernel b) we need to adjust t with respect of n such that ...

% 3) Show experimental results
Define the error, show results and comments

Explain the difference of errors... Uniform sampling is better

>>>>

\subsection{Problem formulation} 
% The plan for this section
% 0.1 Explain the sampling problem
As the graph convolution acts on a sampled version of the signal $\b{f}$ and the rotation on its continuous counterpart $f$, we first need to define a sampling operator.
Given a sampling scheme $V=\{v_i\in\mathbb S^2, i=0, \dots, n\}$ of the sphere and a function $f : \mathbb S^2 \supset F_V \to \mathbb R$, the sampling operator $T_V: L^2(\mathbb S^2) \supset F\to \mathbb R^n,\  T_V(f) = \b{f}$ is defined as $\b{f}:\ f_i=f(v_i)$, where $F_V$ suitable subspace of $L^2(\mathbb S^2)$ such that $T_V$ is invertible, i.e., we can unambiguously reconstruct the function $f\in F_V$ from its sampled values $\b{f}$.
The existence of such subspace depends on the sampling scheme $V$, and its characterization is a common problem in signal processing \cite{Driscoll:1994:CFT:184069.184073}. 
For most samplings, it is not known if $F_V$ exist and hence if $T_V$ is invertible. A special case is the \textit{equiangular sampling} scheme where a sampling theorem holds, and thus we have a closed form of $T_V^{-1}$ is known. % \cite{Driscoll:1994:CFT:184069.184073}
For sampling schemes where no such sampling formula is available, we leverage the discrete SHT to reconstruct $f$ from $\bf{f}$, thus approximating $T_V^{-1}$. 
For all theoretical consideration, we assume that $F_V$ exists and $f \in F_V$.

% 0.2 Explain why the Laplacian is so important...
By definition, our (spherical) graph convolution is rotation equivariant, if and only if, it commutes with the rotation operator defined as $R(g), g\in SO(3)$: $R(g) f(x) = f\left(g^{-1} x \right)$. 
In the context of this work, graph convolution is performed by recursive applications of the graph Laplacian, i.e: $h(\b{L}) \b{f} = \left(\sum_{i=0}^K \alpha_i \b{L}^i\right) \b{f}$\nati{Put this equation somewhere else?}.  Hence, if $\b{L}$ commutes with the rotation operator, then, by recursion, it will also commute with the convolution $h(\b{L})$.
As a result, to simplify our analysis, we concentrate on the graph Laplacian and $h(\b{L})$ is rotational invariant if and only if 
% 0.3 Formulate the problem
\begin{equation} \label{eq:equivariance}
	\b{R}_V(g) \b{L} \b{f} = \b{L} \b{R}_V(g) \b{f} \hspace{1cm} \forall f\in F_V \text{ and } \forall g\in SO(3), 
\end{equation}
where $\b{R}_V(g) = T_V R(g) T_V^{-1}$. For practical application, we define the \textit{normalized equivariance error} for a signal $\b{f}$, and a  rotation $g$ as:
\begin{equation} \label{eq:equivariance error}
	E_{\b{L}}(\b{f}, g) = \left(\frac{ \norm {\b{R}_V(g) \b{L} \b{f} - \b{L} \b{R}_V(g) \b{f}} }{\norm {\b{f}}}\right)^2,
\end{equation}
More generally for a class of signal $f \in C \subset F_V$, the \textit{mean equivariance error}
\begin{equation} \label{eq:mean equivariance error}
	\overline E_{\b{L}, C} = \mathbb E_{\b{f}\in C, g\in SO(3)} \ E_{\b{L}}(\b{f}, g).
\end{equation}
In practice, to anlyse the equivariance error frequency,  the set $C$ we used is composed of functions $f$ with made of spherical harmonic of a single degree $\ell$.
The expected value is obtained by averaging over a finite number of random functions and random rotations.

@@

% 2) explain our approach
We chose a different approach and take inspiration from \cite{belkin2005towards}, which proves that for a random \emph{uniform sampling} scheme the graph Laplacian converges toward the Laplace-Beltrami operator $\Delta_{\mathbb{S}^2}$ as the number of sampling point goes to infinity.
This convergence result is important as a) $\Delta_{\mathbb{S}^2}$ commutes with rotation ($\Delta_{\mathbb{S}^2}R(g) = R(g)\Delta_{\mathbb{S}^2}$) and b) is diagonalized by the spherical harmonics. 
In this case the weighting scheme is full (every node is connected to every node) and is based the exponential kernel: 
\nati{@martino, please check this equation... Do you use 4t?}
\begin{equation} \label{eq:belkin_weights}
w_{ij} = e^{-\frac{1}{t} \norm{x_i-x_j}^2}
\end{equation}
In practice most of the weights are very close to $0$ and hence, to optimize for computational efficiency, we limit ourselves to only the $k$ nearest neighbors. Given $k$, we find the optimal $t$, a grid search is performed using \eqref{eq:mean equivariance error}.

\subsection{Analysis of the proposed weighting scheme}
We analyse our proposed weighting scheme both theoretically and empirically

\paragraph{Theoretical convergence.}
We extend the work Belkin and Nyiogi to a fixed sampling. 
Given a sampling $V = \{x_0, \dots, x_{n-1}\}$ define $\sigma_i$ to be the patch of the surface of the sphere corresponding to $x_i$, define $A_i$ to be its corresponding area and $d_i$ to be the radius of the smallest ball in $\mathbb R^3$ containing $\sigma_i$. Define $d^{(n)} := \max_{i=0, \dots, n}d_i$ and $A^{(n)}=\max_{i=0, \dots, n}A_i$.
\begin{theorem}
	For a sampling $\mathcal P = \{x_i\in\mathbb S^2\}_{i=0}^{n-1}$ of the sphere that is equi area and such that $d^{(n)} \leq \frac{C}{\sqrt{n}}$, for all $f: \mathbb S^2 \rightarrow \mathbb R$ Lipschitz with respect to the euclidean distance in $\mathbb R^3$, for all $y\in\mathbb S^2$, there exists a sequence $t_n = n^\beta$ such that the rescaled Heat Kernel Graph Laplacian operator $\frac{|\mathbb S^2|}{4\pi t_n}L^t_n$ converges pointwise to the Laplace Beltrami operator on the sphere $\triangle_{\mathbb S^2}$  for $n\to\infty$:
	$$ \lim_{n\to\infty} \frac{|\mathbb{S}^2|}{4\pi t_n} \b{L}_n^{t_n} T_{V_n} f(y) =  \Delta_{\mathbb{S}^2}f(y).$$
	\label{theo:pointwise convergence for a regular sampling}
\end{theorem}
While this is \emph{not} a proof of equivariance of the graph convolution for a specific sampling, we observe that the $\b{L}$ converges pointwisely toward an operator ($\Delta_{\mathbb S^2}$) commuting with rotation. 
To have the desired equality in eq..., we would need to prove a stronger result, for example uniform convergence (in the space of the operators).
\nati{Should we put this here?: }
This does not hold in general \cite{belkin2007convergence}. However, we believe that this is achievable for a bandlimited signals.
More importantly, the proof of theorem \ref{theo:pointwise convergence for a regular sampling} in Appendix ... inspires us in the construction of our graph Laplacian. a) use gaussian kernel b) we need to adjust t with respect of n such that ...
\nati{While do not have a proof for it, we srongly believe that the HEALpix sampling do satisfy the hypothesis $d^{(n)}\leq \frac{C}{\sqrt{n}}$ of Theorem \ref{theo:pointwise convergence for a regular sampling}. This is discussed in Appendix~\ref{...}}

% 3) Show experimental results
\paragraph{Empirical convergence.}
\nati{@martino, please put some stuff here...}
Explain the difference of errors... Uniform sampling is better
